---
title: "Case Study"
author: "Zachary Katz"
date: "10/14/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(p8105.datasets)
library(leaflet)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Load a dataset

```{r}
data("nyc_airbnb")
nyc_airbnb
```

* How many rentals are there? Of what type? In what places?
* Is price correlated with review score?
* Which neighborhood is most popular? Most expensive?
* Which neighborhood has the highest review?
* Is availability related to the price?
* Is room type related to neighborhood?
* Is length of stay related to neighborhood (borough)?

```{r}
nyc_airbnb = 
      nyc_airbnb %>% 
      mutate(stars = review_scores_location / 2) %>% 
      rename(borough = neighbourhood_group)
```

Let's try to count things:

```{r}
nyc_airbnb %>% 
      janitor::tabyl(borough, room_type)

# This gives the same thing
nyc_airbnb %>% 
      count(borough, room_type) %>% 
      pivot_wider(
            names_from = room_type,
            values_from = n
      )
```

* Is price correlated with review score?

```{r}

# Boxplots of price vs. stars by borough
nyc_airbnb %>% 
      filter(price < 2500) %>% 
      ggplot(aes(x = as.factor(stars), y = price)) + 
      geom_boxplot() + 
      facet_wrap(. ~ borough)

# Take mean price per star rating and plot
nyc_airbnb %>% 
      group_by(stars) %>% 
      summarize(mean_price = mean(price)) %>% 
      ggplot(aes(x = stars, y = mean_price)) + 
      geom_point() + 
      geom_line() + 
      geom_smooth()

# Table of median price per star rating
nyc_airbnb %>% 
      group_by(stars) %>% 
      summarize(median_price = median(price)) %>% 
      knitr::kable()

# Correlation between star rating and price
cor.test(pull(nyc_airbnb, stars), pull(nyc_airbnb, price))
```

Let's look at price and room type...

```{r}
nyc_airbnb %>% 
      ggplot(aes(x = stars, y = price)) + 
      geom_point() + 
      facet_grid(. ~ room_type)
```

Let's also look at price and neighborhood

```{r}
# Using mean here, but could also do median
nyc_airbnb %>% 
      filter(borough == "Manhattan") %>% 
      group_by(neighbourhood) %>% 
      summarize(mean_price = mean(price, na.rm = TRUE)) %>% 
      arrange(desc(mean_price))

nyc_airbnb %>% 
      filter(borough == "Manhattan", price <= 1000) %>% 
      mutate(neighbourhood = fct_reorder(neighbourhood, price)) %>% 
      ggplot(aes(x = neighbourhood, y = price)) + 
      geom_boxplot() + 
      coord_flip() + 
      facet_grid(. ~ room_type)
      
```

Price vs. location

```{r}
nyc_airbnb %>% 
  filter(price < 500) %>% 
  sample_n(5000) %>% 
  ggplot(aes(x = lat, y = long, color = price)) + 
  geom_point(alpha = 0.5)
```

Note, `sample_n` could to have in back pocket if you know trying to plot everything on one graph is going to break your laptop.

## retry leaflet

```{r}
pal = colorNumeric("viridis", NULL)

nyc_airbnb %>% 
  mutate(
    labels = str_c("<b>Price:", price, "</b<br>Stars:", "stars")
  )

# Couldn't get labels to work, but will try again later
nyc_airbnb %>% 
  filter(price < 500) %>% 
  sample_n(1000) %>% 
  leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addCircleMarkers(~lat, ~long, radius = 1, color = ~pal(price))
```

