---
title: "Exploratory data analysis"
output: github_document
---

Load in a dataset that we'll use often.

```{r}
# Load packages
library(tidyverse)
library(viridis)
library(ggridges)
library(patchwork)

# Set global options for embedding plots
knitr::opts_chunk$set(
      fig.width = 6,
      fig.asp = 0.6,
      out.width = "90%"
)

# Get data from NOAA and clean it up a bit

weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
## Registered S3 method overwritten by 'hoardr':
##   method           from
##   print.cache_info httr
## using cached file: /Users/jeffgoldsmith/Library/Caches/R/noaa_ghcnd/USW00094728.dly
## date created (size, mb): 2020-09-25 14:56:47 (7.519)
## file min/max dates: 1869-01-01 / 2020-09-30
## using cached file: /Users/jeffgoldsmith/Library/Caches/R/noaa_ghcnd/USC00519397.dly
## date created (size, mb): 2020-09-25 14:56:52 (1.699)
## file min/max dates: 1965-01-01 / 2020-03-31
## using cached file: /Users/jeffgoldsmith/Library/Caches/R/noaa_ghcnd/USS0023B17S.dly
## date created (size, mb): 2020-09-25 14:56:54 (0.877)
## file min/max dates: 1999-09-01 / 2020-09-30
```

## `group_by`

```{r}
weather_df %>% 
  group_by(name, month)
```

## count some things

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n())

# Shortcut way
weather_df %>% 
  count(name, month)
```

It's nice that summarize produces a data frame

`table` doesn't, and that's bad.

```{r}
weather_df %>% 
  pull(name) %>% 
  table()
```

```{r}
weather_df %>% 
  janitor::tabyl(month, name)
```

## More general summaries

Let's count, take means, look at SDs...

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    n_obs = n(),
    # Could use na.rm, or drop_na earlier in tidying
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmax = median(tmax, na.rm = TRUE),
    sd_tmax = sd(tmax, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
  geom_point() + 
  geom_line()
```

Formatting table outputs...

```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  knitr::kable(digits = 2)
```

## grouped `mutate`

```{r}
# Adding a column to the data frame in a way that is group specific
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax
  ) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
  geom_point() +
  geom_smooth()
```


```{r}
# Coldest day in each place, use <2 or >2 because of ties
weather_df %>% 
  group_by(name) %>% 
  mutate(
    tmax_rank = min_rank(tmax)
  ) %>% 
  filter(tmax_rank < 2)

# Warmest day in each place, use <2 or >2 because of ties
weather_df %>% 
  group_by(name) %>% 
  mutate(
    tmax_rank = min_rank(desc(tmax))
  ) %>% 
  filter(tmax_rank < 2)
```

Note that different functions deal with ties in different ways.

Lagged variables:

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    # What is the tmax observation I had the day before?
    # To construct lag information, you should have the grouping structure in place, otherwise will inadvertently put things in the wrong order
    # Order that the dataframe is in makes a difference (e.g. dates arranged as they need to be for `lag`)
    lagged_tmax = lag(tmax, n = 1),
    tmax_diff = tmax - lagged_tmax
  ) %>% 
  summarize(diff_sd = sd(tmax_diff, na.rm = TRUE))
```

## Limitations

What if my "summary" is a linear model ...

```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(cor_tmin_tmax = cor(tmin, tmax, use = "complete"))

weather_df %>% 
  # Cannot replace this with group_by and then try to run linear models on each place, since in one data frame
  filter(name == "CentralPark_NY") %>% 
  lm(tmax ~ tmin, data = .)
```

```{r eval = FALSE}
# This doesn't work because lm must be a vector, not an lm object
weather_df %>% 
  group_by(name) %>% 
  summarize(lm = lm(tmax ~ tmin))
```


